---
title: "Introduction to BayesGmed"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BayesGmed}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Motivation 
This vignette introduces the R package *BayesGmed*, a package designed for a Bayesian causal mediation analysis in Stan, a probabilistic programming language for Bayesian inference. 
BayesGmed uses a parametric approach for the specification of the outcome and mediator model, and uses the g-formula approach for estimation. In addition to the estimation of causal mediation effects, the package also allows researchers to conduct sensitivity analysis.

# Getting started 
You can install from Github via devtools
```{r installation, eval= FALSE}
devtools::install_gitlab(belayb/BayesGmed)
```

# Example 
*BayesGmed* comes with an example data, example_data, 
which contains a binary outcome $Y$, a single binary mediator $M$, a binary exposure $A$, 
and two numeric confounding variables $Z = (Z_1, Z_2)$. The first 6 rows of the data is
visualized below. 
```{r setup}
library(BayesGmed)
```
```{r data}
head(example_data)
```
The aim in this example data is to estimate the direct and indirect effect of $A$ on $Y$ adjusting for $Z$. 
To do this, we may proced as follow. 

\[logit( P(Y_{i} = 1|A_{i},M_{i},\mathbf{Z}_{i}) ) = \alpha_{0} + \mathbf{\alpha}_{Z}^{'}\mathbf{Z}_{i} + \alpha_{A}A_{i} + \alpha_{M}M_{i},\]

\[logit(P(M_{i} = 1| A_{i},\mathbf{Z}_{i} ) ) = \beta_{0} + \mathbf{\beta}_{Z}^{'}\mathbf{Z}_{i} + \beta_{A}A_{i}.\]

To complete the model specification, we assume the following priors for the model parameters. 

$$
\begin{aligned}
\beta &\sim MVN(location_m, scale_m)\\
\alpha &\sim MVN(location_y, scale_y)
\end{aligned}
$$

We then need to specify the parameters 
of the prior distrbutions or assume a hyper-prior. 
BayesGmed currently only allow fixed values and these values
can be passed as part of the model call and if not the 
following defult values will be used 

```{r prior, eval= FALSE}

priors <- list(scale_m = 2.5*diag(P+1), 
               scale_y = 2.5*diag(P+2),
               location_m = rep(0, P+1), 
               location_y = rep(0, P+2))
```

The model can then be fitted as follow 

```{r model}
fit1 <- bayesgmed(outcome = "Y", mediator =  "M", treat = "A", covariates = c("Z1", "Z2"), dist.y = "binary",
dist.m = "binary", link.y = "logit", link.m = "logit", data = example_data)

bayesgmed_summary(fit1)
```